---
title: "TimeSeries"
author: "Sergio Solano"
date: "9 de marzo de 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library("astsa")
library("xts")
library ("vars")
library ("tsDyn")
```
## Pruebas series de tiempo de Time Series Analysis and its applications

```{r Datacamp Funciones, include=TRUE, echo = TRUE}
# Plot AirPassengers
plot(AirPassengers)

# View the start and end dates of AirPassengers
start(AirPassengers)
end(AirPassengers)

# Use time(), deltat(), frequency(), and cycle() with AirPassengers 
time(AirPassengers)
deltat(AirPassengers)
frequency(AirPassengers)
cycle(AirPassengers)

# Simulate a WN model with list(order = c(0, 0, 0))
white_noise <- arima.sim(model = list(order = c(0,0,0)), n = 50)

# Plot your white_noise data

ts.plot(white_noise)

# Simulate from the WN model with: mean = 100, sd = 10
white_noise_2 <- arima.sim(model = list(order = c(0,0,0)), n = 50, mean = 100, sd = 10)

# Plot your white_noise_2 data
ts.plot(white_noise_2)


```
Los cambios en una serie de tiempo de Random Walk siguen un comportamiento de White noise. 

```{r white noise}
# Generate a RW model using arima.sim
random_walk <- arima.sim(model = list(order = c(0, 1, 0)) , n = 100)

# Plot random_walk
ts.plot (random_walk)

# Calculate the first difference series
random_walk_diff <- diff(random_walk) 

# Plot random_walk_diff

ts.plot(random_walk_diff)

# RANDOM WALK WITH DRIFT

# Generate a RW model with a drift uing arima.sim
rw_drift <- arima.sim(model = list(order = c(0, 1, 0)), n = 100, mean = 1)

# Plot rw_drift
ts.plot(rw_drift)

# Calculate the first difference series
rw_drift_diff <- diff(rw_drift) 

# Plot rw_drift_diff
ts.plot(rw_drift_diff)


random_walk <- rw_drift

# Difference your random_walk data
rw_diff <- diff(random_walk)


# Plot rw_diff
plot.ts(rw_diff)

# Now fit the WN model to the differenced data
model_wn <- arima(rw_diff,order = c(0, 0, 0))

# Store the value of the estimated time trend (intercept)
int_wn <- model_wn$coef

# Plot the original random_walk data
ts.plot(random_walk)

# Use abline(0, ...) to add time trend to the figure
abline(0,int_wn)

#  RANDOM WALK CON Y SIN DRIFT (Estationarity)
# 
# The white noise (WN) and random walk (RW) models are very closely related. However, only the RW is always non-stationary, both with and without a drift term. This is a simulation exercise to highlight the differences.
# 
# Recall that if we start with a mean zero WN process and compute its running or cumulative sum, the result is a RW process. The cumsum() function will make this transformation for you. Similarly, if we create a WN process, but change its mean from zero, and then compute its cumulative sum, the result is a RW process with a drift.

# Use arima.sim() to generate WN data
white_noise <- arima.sim(model = list(order = c(0, 0, 0)), n = 100) 
 
# Use cumsum() to convert your WN data to RW
random_walk <- cumsum(white_noise)
  
# Use arima.sim() to generate WN drift data
wn_drift <- arima.sim(model = list(order = c(0, 0, 0)),mean=0.4, n = 100)
  
# Use cumsum() to convert your WN drift data to RW
rw_drift <- cumsum(wn_drift)

# Plot all four data objects
plot.ts(cbind(white_noise, random_walk, wn_drift, rw_drift))

# 
# # Generate means from eu_percentreturns
# colMeans(eu_percentreturns)
# 
# # Use apply to calculate sample variance from eu_percentreturns
# apply(eu_percentreturns, MARGIN = 2, FUN = var)
# 
# # Use apply to calculate standard deviation from eu_percentreturns
# apply(eu_percentreturns, MARGIN = 2, FUN = sd)
# 
# # Display a histogram of percent returns for each index
# par(mfrow = c(2,2))
# apply(eu_percentreturns, MARGIN = 2, FUN = hist, main = "", xlab = "Percentage Return")
# 
# # Display normal quantile plots of percent returns for each index
# par(mfrow = c(2,2))
# apply(eu_percentreturns, MARGIN = 2, FUN = qqnorm, main = "")
# qqline(eu_percentreturns)

# pairs(eu_stocks)


# MODELOS AUTOREGRESIVOS: 

# Simulate an AR model with 0.5 slope
x <- arima.sim(model = list(ar = 0.5), n = 100)

# Simulate an AR model with 0.9 slope
y <- arima.sim(model = list(ar = 0.9), n = 100)

# Simulate an AR model with -0.75 slope
z <- arima.sim(model = list(ar = -0.75), n = 100)

# Plot your simulated data
plot.ts(cbind(x, y, z))

###### COMPARAR MODELOS AR con RandomWalks 

# Simulate and plot AR model with slope 0.9 
x <- arima.sim(model = list(ar = 0.9), n = 200)
ts.plot(x)
acf(x)

# Simulate and plot AR model with slope 0.98
y <- arima.sim(model = list(ar = 0.98), n = 200)
ts.plot(y)
acf(y)

# Simulate and plot RW model
z <- arima.sim(model = list(order = c(0, 1, 0)), n = 200)
ts.plot(z)
acf(z)

# # AJUSTAR UN AR
# # Fit an AR model to Nile
# AR_fit <-arima(Nile, order  = c(1,0,0))
# print(AR_fit)
# 
# # Use predict() to make a 1-step forecast
# predict_AR <- predict(AR_fit)
# 
# # Obtain the 1-step forecast using $pred[1]
# predict_AR$pred[1]
# 
# # Use predict to make 1-step through 10-step forecasts
# predict(AR_fit, n.ahead = 10)
# 
# # Run to plot the Nile series plus the forecast and 95% prediction intervals
# ts.plot(Nile, xlim = c(1871, 1980))
# AR_forecast <- predict(AR_fit, n.ahead = 10)$pred
# AR_forecast_se <- predict(AR_fit, n.ahead = 10)$se
# points(AR_forecast, type = "l", col = 2)
# points(AR_forecast - 2*AR_forecast_se, type = "l", col = 2, lty = 2)
# points(AR_forecast + 2*AR_forecast_se, type = "l", col = 2, lty = 2)

# # AJUSTAR UN MA
# Generate MA model with slope 0.5
x <- arima.sim(model = list(ma = 0.5), n = 100)

# Generate MA model with slope 0.9
y <- x <- arima.sim(model = list(ma = 0.9), n = 100)

# Generate MA model with slope -0.5
z <- x <- arima.sim(model = list(ma = -0.5), n = 100)

# Plot all three models together
plot.ts(cbind(x, y, z))
# # # Ajustando un MA  
# 
# # Make a 1-step forecast based on MA
# predict_MA <- predict(MA, n.ahead = 1)
# 
# # Obtain the 1-step forecast using $pred[1]
# predict_MA$pred[1] 
# 
# # Make a 1-step through 10-step forecast based on MA
# predict(MA, n.ahead = 10)
# 
# # Plot the Nile series plus the forecast and 95% prediction intervals
# ts.plot(Nile, xlim = c(1871, 1980))
# MA_forecasts <- predict(MA, n.ahead = 10)$pred
# MA_forecast_se <- predict(MA, n.ahead = 10)$se
# points(MA_forecasts, type = "l", col = 2)
# points(MA_forecasts - 2*MA_forecast_se, type = "l", col = 2, lty = 2)
# points(MA_forecasts + 2*MA_forecast_se, type = "l", col = 2, lty = 2)

```


## Pruebas series de tiempo de Time Series Analysis and its applications course

```{r curso ARIMA}
# # R Arima base 

# View a detailed description of AirPassengers
help(AirPassengers)

# Plot AirPassengers
ts.plot(AirPassengers)

# Plot the DJIA daily closings
ts.plot(djia$Close)

# Plot the Southern Oscillation Index
plot(soi)

# Plot globtemp and detrended globtemp
par(mfrow = c(2,1))
plot(globtemp) 
plot(diff(globtemp))

# Plot cmort and detrended cmort
par(mfrow = c(2,1))
plot(cmort)
plot(diff(cmort))


# Plot GNP series (gnp) and its growth rate
par(mfrow = c(2,1))
plot(gnp)
plot(diff(log(gnp)))


# Plot DJIA closings (djia$Close) and its returns
par(mfrow = c(2,1))
plot(djia$Close)
plot(diff(log(djia$Close)))

##SIMULAR MA & AR

# Generate and plot white noise
WN <- arima.sim(model=list(order=c(0,0,0)),n=100)
plot(WN)
# Generate and plot an MA(1) with parameter .9 
MA <- arima.sim(model=list(order=c(0,0,1), ma = 0.9),n=100)
plot(MA)

# Generate and plot an AR(2) with parameters 1.5 and -.75
AR <- arima.sim(model=list(order=c(2,0,0), ar =c(1.5,-0.75)),n=100)
plot(AR)

# # SIMULAR Y AJUSTAR AR


# Generate 100 observations from the AR(1) model
x <- arima.sim(model = list(order = c(1, 0, 0), ar = .9), n = 100) 

# Plot the generated data 
plot(x)

# Plot the sample P/ACF pair
acf2(x)

# Fit an AR(1) to the data and examine the -table
sarima(x, 1, 0, 0)

# # aJUSTAR MA(2)

# astsa is preloaded
x <- arima.sim(model = list(order = c(2, 0, 0), ar = c(1.5, -.75)), n = 200)
# Plot x
plot(x)

# Plot the sample P/ACF of x
acf2(x)

# Fit an AR(2) to the data and examine the t-table
sarima(x,2,0,0)

# # # SIMULANDO ARIMAS 

# Fit an ARIMA(0,1,2) to globtemp and check the fit
sarima(globtemp, 0, 1, 2)

# Forecast data 35 years into the future
sarima.for (globtemp,n.ahead = 35, p = 0, d = 1, q = 2) 

### SERIES DE TIEMPO ESTACIONALES 

# Plot unemp 
plot(unemp)

# Difference your data and plot it
d_unemp <- diff(unemp) 
plot(d_unemp)

# Seasonally difference d_unemp and plot it
dd_unemp <- diff(d_unemp, lag = 12)  
plot(dd_unemp)

# Plot P/ACF pair of fully differenced data to lag 60
dd_unemp <- diff(diff(unemp), lag = 12)
acf2(dd_unemp, max.lag = 60)

# Fit an appropriate model
sarima(unemp,2,1,0,0,1,1, S=12)

# Plot differenced chicken
plot(diff(chicken))

# Plot P/ACF pair of differenced data to lag 60
acf2(diff(chicken), max.lag = 60)

# Fit ARIMA(2,1,0) to chicken - not so good
sarima(chicken, 2,1,0)

# Fit SARIMA(2,1,0,1,0,0,12) to chicken - that works
sarima(chicken,2,1,0,1,0,0,12)

###### Natalidad US 

# Plot P/ACF to lag 60 of differenced data
d_birth <- diff(birth)
acf2(d_birth)

# Plot P/ACF to lag 60 of seasonal differenced data
dd_birth <- diff(d_birth, lag = 12)
acf2(dd_birth)
# Fit SARIMA(0,1,1)x(0,1,1)_12. What happens?
sarima(birth,0,1,1,0,1,1,12)

# Add AR term and conclude
sarima(birth,1,1,1,0,1,1,12)

## FORECASTING SEASONAL DATA 

# Fit your previous model to unemp and check the diagnostics
sarima(unemp, 2,1,0,0,1,1,12)

# Forecast the data 3 years into the future
sarima.for(unemp,n.ahead = 36,2,1,0,0,1,1,12)

# Fit the chicken model again and check diagnostics
sarima(chicken,2,1,0,1,0,0,12)

# Forecast the chicken data 5 years into the future
sarima.for(chicken,n.ahead = 60,2,1,0,1,0,0,12)


```



## Pruebas series de tiempo de Time Series Analysis and its applications text 


Página 13
Moving Average

```{r cars}
w = rnorm (500,0,1)
v = filter(w, sides = 2, rep(1/3,3))
par(mfrow=c(2,1))
plot.ts(w,main = "white noise")
plot.ts(v, main = "moving average")
```

Autoregesivo

```{r}
w = rnorm(550,0,1)
x = filter (w, filter = c(1,-0.9), method = "recursive")[-(1:50)]
plot.ts(x, main = "autoregression")

```

Autoregesivo con drift 

```{r}
set.seed(154)
w = rnorm(200,0,1); x = cumsum (w)
wd = w + .2; xd = cumsum(wd)
plot.ts(xd, ylim = c(-5,55), main = "random walk")
lines(x); lines(0.2*(1:200), lty = "dashed")

```

ACF

```{r}
par(mfrow = c(3,1))
acf(soi, 48, main = "Southern Oscilation Index")
acf(speech, 250)

par(mfrow=c(3,1))
acf(soi, 48, main="Southern Oscillation Index") 
acf(rec, 48, main="Recruitment")
ccf(soi, rec, 48, main="SOI vs Recruitment", ylab="CCF")

# Varios vectores

persp(1:64, 1:36, soiltemp, phi=30, theta=30, scale=FALSE, expand=4, ticktype="detailed", xlab="rows", ylab="cols", zlab="temperature")

plot.ts(rowMeans(soiltemp), xlab="row", ylab="Average Temperature")


```

# VECTORES AUTOREGRESIVOS 

```{r vectores autoregresivos}

B1<-matrix(c(0.7, 0.2, 0.2, 0.7), 2)
var1<-VAR.sim(B=B1,n=100,include="none")
ts.plot(var1, type="l", col=c(1,2))
VARselect(var1, lag.max=9, type="const")
var1_est <- VAR(var1,p = 1)
var1_est
```

#MODELOS ESTADO ESPACIO
El paquete base de R permite evaluar modelos estado espacio con la función [link](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/Nile.html) StructTS

```{r}
# rm(list = ls())
fitNile <- StructTS(Nile, "level")
fitNile
tsdiag(fitNile)
plot(Nile, type = "l") 
lines(fitted(fitNile), lty = "dashed", lwd = 2) 
lines(tsSmooth(fitNile), lty = "dotted", lwd = 2)

SP_fit <- StructTS(Nile, "level")
SP_forecast <- predict(SP_fit, n.ahead = 100)$pred
SP_forecast_se <- predict(SP_fit, n.ahead = 100)$se
plot(Nile, type = "l") 
points(SP_forecast, type = "l", col = 2)
points(SP_forecast - 2*SP_forecast_se, type = "l", col = 2, lty = 2)
points(SP_forecast + 2*SP_forecast_se, type = "l", col = 2, lty = 2)


```




